---
title: "Chicago's 77 Neighborhoods"
subtitle: "Business activity and community sentiment — a data portrait"
author: "Data: Chicago Data Portal + Reddit"
date: last-modified
format:
  revealjs:
    theme: [default, custom.scss]
    slide-number: true
    hash: true
    transition: slide
    controls: true
    embed-resources: true
    width: 1280
    height: 720
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false

library(here)
library(dplyr)
library(ggplot2)
library(leaflet)
library(sf)
library(scales)
library(knitr)
library(stringr)
library(purrr)

library(patchwork)
source(here("R", "utils", "chart_functions.R"))

# Load master dataset
metrics_path <- here("data", "processed", "neighborhood_metrics.rds")
if (!file.exists(metrics_path)) {
  stop("neighborhood_metrics.rds not found. Run the pipeline first:\n",
       "  Rscript run_pipeline.R")
}

neighborhood_metrics <- readRDS(metrics_path)

# Load TF-IDF terms for neighborhood charts
tfidf_data <- readRDS(here("data", "processed", "tfidf_terms.rds"))

# Load Airbnb TF-IDF terms (visitor voice)
airbnb_tfidf_path <- here("data", "processed", "airbnb_tfidf.rds")
airbnb_tfidf_data <- if (file.exists(airbnb_tfidf_path)) {
  readRDS(airbnb_tfidf_path)
} else {
  tibble(slug = character(), word = character(), tf_idf = numeric())
}

# Load business category breakdown for charts
biz_data       <- readRDS(here("data", "processed", "business_summary.rds"))
biz_by_category <- biz_data$by_category

# Optional: subset for testing
subset_n <- as.integer(Sys.getenv("NEIGHBORHOOD_SUBSET", unset = "0"))
if (subset_n > 0 && subset_n < 77) {
  message("NEIGHBORHOOD_SUBSET=", subset_n, " — rendering subset of ", subset_n, " neighborhoods")
  neighborhood_metrics <- head(neighborhood_metrics, subset_n)
}

geojson_path <- here("data", "reference", "chicago_community_areas.geojson")
```

## Chicago's 77 Community Areas {.title-slide}

:::{.columns}
:::{.column width="60%"}
An automated data portrait of every official Chicago neighborhood, combining:

- **Active business licenses** (Chicago Data Portal)
- **Resident voice** (Reddit discussions)
- **Visitor voice** (Airbnb guest reviews)
- **AI-generated summaries** (Claude claude-haiku-4-5-20251001)

`r nrow(neighborhood_metrics)` neighborhoods · `r scales::comma(sum(neighborhood_metrics$total_active_businesses))` active businesses

:::
:::{.column width="40%"}

```{r sentiment-summary}
sent_counts <- count(neighborhood_metrics, sentiment_label)
knitr::kable(
  sent_counts,
  col.names = c("Sentiment", "Neighborhoods"),
  align = c("l", "r")
)
```

:::
:::

---

## Neighborhood Sentiment Map

```{r choropleth}
#| out-width: "100%"
#| out-height: "550px"
if (file.exists(geojson_path)) {
  make_overview_choropleth(geojson_path, neighborhood_metrics)
} else {
  cat("GeoJSON not found. Download from Chicago Data Portal and save to data/reference/")
}
```

*Colored by normalized AFINN sentiment score (green = positive, red = negative, grey = no Reddit data)*

---

## By the Numbers

```{r overview-stats}
#| fig-height: 4
top10 <- neighborhood_metrics |>
  slice_max(order_by = total_active_businesses, n = 10) |>
  mutate(name = forcats::fct_reorder(name, total_active_businesses))

ggplot(top10, aes(x = total_active_businesses, y = name, fill = sentiment_label)) +
  geom_col(width = 0.65) +
  geom_text(aes(label = comma(total_active_businesses)), hjust = -0.15, size = 3.2,
            family = "serif", color = "#4a3f35") +
  scale_fill_manual(
    values = c("Positive" = "#4a7c59", "Neutral" = "#b8963e", "Negative" = "#9e3a3a"),
    name   = "Sentiment"
  ) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15)), labels = comma) +
  labs(
    title = "Top 10 Neighborhoods by Active Business Count",
    x     = "Active business licenses",
    y     = NULL
  ) +
  theme_tufte_slide(base_size = 13)
```

---

```{r neighborhood-slides, results='asis'}
# Loop over each neighborhood and knit the child slide template.
# The child reads `current_slug` from the environment.

slugs_to_render <- neighborhood_metrics$slug

for (current_slug in slugs_to_render) {
  # Make current_slug available to child document
  knitr::knit_child(
    here("slides", "_neighborhood_slide.qmd"),
    envir = environment(),
    quiet = TRUE
  ) |> cat()
}
```

---

## Data & Methods

:::{.columns}
:::{.column width="50%"}

**Business Data**

- Source: Chicago Data Portal Socrata API
- Filter: `license_status = AAC` (active)
- ~200 license types → 10 categories
- Diversity: Shannon entropy

**Reddit Data** (resident voice)

- Public JSON API (no OAuth)
- r/chicago + neighborhood subreddits
- Past 12 months
- Sentiment: AFINN + Bing lexicons (tidytext)

**Airbnb Data** (visitor voice)

- Inside Airbnb (insideairbnb.com), 2 snapshots
- Spatial join to community areas via lat/lon
- NLP on guest review text: bi/tri-gram TF-IDF

:::
:::{.column width="50%"}

**AI Summaries**

- Model: `claude-haiku-4-5-20251001`
- Synthesizes business + resident + visitor data
- Flags divergence between Reddit and Airbnb themes
- Cached; not regenerated each render

**Data Quality Flags**

- `full` — business + Reddit (≥10 posts)
- `low_reddit` — <10 Reddit posts
- `business_only` — no Reddit data
- `no_airbnb` — no Airbnb listings found

**Code:** github.com/[your-username]/chicago_neighborhoods

:::
:::
